Paper figures

### 2019-01-07 paper_figures.Rmd
new panel

---
title: "Paper figures"
author: "Francesco Beghini"
date: "January 6, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

source('R/loadQiimeData.R')
source('R/ggtheme.R')
```

```{r QIIME data loading}
gg_topi <- loadQiimeData()
gg_topi.genus <- tax_glom(gg_topi,taxrank = "Genus")
gg_topi.genus.relab <- transform_sample_counts(gg_topi.genus, function(x) x/sum(x))

@@

# Figure 3
## Panel A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3 
```{r Calculate alpha diversity}
alpha_div <- estimate_richness(gg_topi) %>% bind_cols(data.frame(sample_data(gg_topi)))
```

@@

## Panel B: Beta diversity differences pre and post challange T2 vs T3
```{r}
distwu <- phyloseq::distance(gg_topi, "wunifrac")
ordwu <-  ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
distanceswu <- calculate_pairwise_distances(distwu)
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T2','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T2','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)
```

@@

subset(ordwu, timepoint %in% c('T7','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=timepoint), size=3, alpha=.8) +
  scale_color_brewer(name = "Timepoint", palette="Dark2") + 
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(bquote(atop("p-value"==.(permanova$aov.tab$`Pr(>F)`[1]),R^2==.(signif(permanova$aov.tab$R2[1], 2)))), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0, xmax = 0, ymin = .1, ymax = .1)
```

@@

otu_table(gg_topi.genus.relab)[c(idx_247,idx_lachno),] %>%
  data.frame %>%
  tibble::rownames_to_column("OTU") %>%
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
  bind_cols(data.frame(tax_table(gg_topi.genus.relab)[.$OTU,c('Phylum','Family','Genus')])) %>%
  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus))) %>%
  ggplot() +
  geom_boxplot(aes(timepoint, Abundance, fill=condition)) +
  scale_fill_brewer(name = "Timepoint", palette="Dark2") +
  facet_grid(taxlab ~ ., scales = "free", space = "free_x", as.table= T) +
  theme_cm()
```

>>>>

---
title: "Paper figures"
author: "Francesco Beghini"
date: "January 6, 2019"
output: 
  html_document: 
    fig_caption: yes
editor_options: 
  chunk_output_type: console
---

@@

source('R/loadQiimeData.R')
source('R/ggtheme.R')
source('R/calculatePairwiseDistances.R')
```

```{r QIIME data loading, cache=TRUE}
gg_topi <- loadQiimeData()
mappingfile <- data.frame(sample_data(gg_topi))
gg_topi.genus <- tax_glom(gg_topi,taxrank = "Genus")
gg_topi.genus.relab <- transform_sample_counts(gg_topi.genus, function(x) x/sum(x))

@@

# Figure 3
## Panel A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3 
```{r Calculate alpha diversity, cache=TRUE}
alpha_div <- estimate_richness(gg_topi) %>% bind_cols(data.frame(sample_data(gg_topi)))
```

@@

## Panel B: Beta diversity differences pre and post challange T2 vs T3
```{r, cache=TRUE}
distwu <- phyloseq::distance(gg_topi, "wunifrac")
ordwu <-  ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
distanceswu <- calculate_pairwise_distances(distwu)
distancesbc <- calculate_pairwise_distances(phyloseq::distance(gg_topi, "bray"))
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T2','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T2','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)
```

@@

subset(ordwu, timepoint %in% c('T7','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=timepoint), size=3, alpha=.8) +
  scale_color_brewer(name = "Timepoint", palette="Dark2") + 
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(bquote(atop(italic("p-value")==.(permanova$aov.tab$`Pr(>F)`[1]),italic(R^2)==.(signif(permanova$aov.tab$R2[1], 2)))), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0, xmax = 0, ymin = .1, ymax = .1)
```

@@

otu_table(gg_topi.genus.relab)[c(idx_247,idx_lachno),] %>%
  data.frame %>%
  tibble::rownames_to_column("OTU") %>%
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
  bind_cols(data.frame(tax_table(gg_topi.genus.relab)[.$OTU,c('Phylum','Family','Genus')])) %>%
  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus))) %>%
  ggplot() +
  geom_boxplot(aes(timepoint, Abundance, fill=condition)) +
  scale_fill_manual(name = "Treatment", values = c("#a6d854","#ffd92f","#66c2a5","#fc8d62")) +
  facet_wrap(facets = 'taxlab', nrow = 2, scales = "free", shrink = FALSE) +
  theme_cm() +
  ylab('Relative abundance') +
  xlab('Timepoint')
```

## Panel G: Scatterplot of tumor size and difference of \\beta diversity 
```{r}
brayc_ds <- distancesbc %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))

ds_corr_braycf <- data.frame()
for(x in unique(brayc_ds$timepoint2)){
  # for(y in levels(brayc_ds$condizione)){
    tmp_res <- with(brayc_ds %>% filter(timepoint2==x), cor.test(value, tumor_mass))
    ds_corr_braycf %<>% bind_rows(data.frame(timepoint = paste('T', x, '-T3',sep = ''), corr = signif(tmp_res$estimate,2), pval = signif(tmp_res$p.value,2), label = paste('p-value==',signif(tmp_res$p.value,2),'~italic(r)==', signif(tmp_res$estimate,2))))
}
```

```{r corr tumor braycurtis}
distancesbc %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  mutate(timepoint = paste('T',timepoint2,'-T3',sep = '')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() +
  facet_wrap(facets = 'timepoint', ncol = 1) +
  geom_line(stat='smooth', method = 'lm', color='red', alpha=0.7, size=.7) + 
  ylab(bquote('Tumor mass'~( mm^3))) +
  xlab('Bray Curtis distance') +
  theme_cm() +
  geom_text(data = ds_corr_braycf, aes(x=0.4,y=2500, label=label), parse = TRUE, inherit.aes = FALSE)
